---
title: "GMBA SWE anomalies 2024"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
    theme: 
      version: 5
      bootswatch: morph
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(stringr)
library(sf)
library(leaflet)
library(htmltools)
library(rmapshaper)
```


```{r data-prep-era5l}
sf_swe <- read_sf("data-raw/SWE 1st Apr anomaly 2024.geojson")

# sf_swe2 <- st_simplify(sf_swe, preserveTopology = T, dTolerance = 2000)
# sf_swe3 <- st_cast(sf_swe2, "MULTIPOLYGON")

sf_swe2 <- ms_simplify(sf_swe, keep = 0.2, keep_shapes = FALSE)

sf_plot <- sf_swe2 %>% 
  filter(!is.na(relAno2024)) %>% 
  mutate(xx_col = if_else(relAno2024 > 150, 150, relAno2024) %>% as.integer) %>% 
  rename(Name = MapName, 
         peak_SWE_2024_anomaly = relAno2024,
         peak_SWE_2024 = peakSweCurrent20242024_mean,
         peak_SWE_clim_1992_2020 = peakSwe19912020_mean) %>% 
  mutate(labels = str_c(
    "<table>",
    "<tr><th style='text-align: left'>", Name, "</th></tr>",
    "<tr>", 
    "<td>", "01 April SWE 2024 anomaly:", "</td>",
    "<td style='text-align: right; padding-left: 15px;'>", sprintf("%i %%", peak_SWE_2024_anomaly), "</td>",
    "</tr>",
    "<tr>", 
    "<td>", "01 April SWE 2024:", "</td>",
    "<td style='text-align: right; padding-left: 15px;'>", sprintf("%0.1f mm", 1000*peak_SWE_2024), "</td>",
    "</tr>",
    "<tr>", 
    "<td>", "01 April SWE average 1991-2020:", "</td>",
    "<td style='text-align: right; padding-left: 15px;'>", sprintf("%0.1f mm", 1000*peak_SWE_clim_1992_2020), "</td>",
    "</tr>",
    "</table>"
  )) 

# pal <- colorNumeric("RdYlBu", domain = sf_plot$xx_col, reverse = F)
# pal2 <- colorNumeric("RdYlBu", domain = sf_plot$xx_col, reverse = T)

pal_pos <- colorRampPalette(colors = c("#ffffbf", "#2c7bb6"), space = "Lab")(150)
pal_neg <- colorRampPalette(colors = c("#d7191c", "#ffffbf"), space = "Lab")(100)

pal <- colorNumeric(palette = c(pal_neg, pal_pos), domain = c(-100, 150))
pal2 <- colorNumeric(palette = c(pal_neg, pal_pos), domain = c(-100, 150), reverse = T)

# previewColors(colorNumeric(palette = c(pal_neg, pal_pos), domain = NULL), values = -100:150)

```


```{r data-prep-stations}

sf_snotel <- readr::read_csv("data-raw/1Apil2024_SNOTEL_anom.csv",
                             show_col_types = FALSE) %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326)

sf_bcab <- readr::read_csv("data-raw/1Apil2024_BCAB_anom.csv",
                           show_col_types = FALSE) %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326)

sf_stn <- bind_rows(sf_snotel, sf_bcab)

sf_plot_stn <- sf_stn %>% 
  mutate(xx_col = if_else(anom_pct > 150, 150, anom_pct) %>% as.integer) %>% 
  mutate(Name = str_to_title(station_name)) %>% 
  rename(peak_SWE_2024_anomaly = anom_pct,
         peak_SWE_2024 = snw,
         peak_SWE_clim_1992_2020 = clm) %>% 
  mutate(labels = str_c(
    "<table>",
    "<tr><th style='text-align: left'>", Name, "</th></tr>",
    "<tr>", 
    "<td>", "01 April SWE 2024 anomaly:", "</td>",
    "<td style='text-align: right; padding-left: 15px;'>", sprintf("%i %%", peak_SWE_2024_anomaly), "</td>",
    "</tr>",
    "<tr>", 
    "<td>", "01 April SWE 2024:", "</td>",
    "<td style='text-align: right; padding-left: 15px;'>", sprintf("%0.1f mm", peak_SWE_2024), "</td>",
    "</tr>",
    "<tr>", 
    "<td>", "01 April SWE average 1991-2020:", "</td>",
    "<td style='text-align: right; padding-left: 15px;'>", sprintf("%0.1f mm", peak_SWE_clim_1992_2020), "</td>",
    "</tr>",
    "</table>"
  )) 

```



# Map

## Column 

### Map

```{r map}
leaflet(sf_plot) %>%
  addProviderTiles("CartoDB.Positron", group = "CartoDB") %>% 
  addProviderTiles("Esri.WorldTopoMap", group = "Topomap") %>% 
  addProviderTiles("Esri.WorldImagery", group = "WorldImagery") %>% 
  addPolygons(
    data = sf_plot,
    fillColor = ~pal(xx_col),
    fillOpacity = 1,
    color = ~pal(xx_col),
    opacity = 1,
    # stroke = FALSE,
    weight = 0,
    group = "ERA5-Land",
    highlightOptions = highlightOptions(
      color = "black",
      weight = 2,
      # dashArray = ""
      bringToFront = FALSE
    ),
    label = lapply(sf_plot$labels, htmltools::HTML),
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "12px",
      offset = c(20, 20),
      direction = "auto")
  ) %>%
  addCircleMarkers(
    data = sf_plot_stn,
    stroke = TRUE,
    weight = 1,
    color = grey(0.3),
    opacity = 1,
    fillOpacity = 1,
    fillColor = ~pal(xx_col),
    radius = 5,
    group = "In-situ",
    label = lapply(sf_plot_stn$labels, htmltools::HTML),
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "12px",
      offset = c(20, 20),
      direction = "auto")
  ) %>% 
  addLayersControl(
    baseGroups = c("CartoDB", "Topomap", "WorldImagery"),
    overlayGroups = c("ERA5-Land", "In-situ"),
    position = "topright"
  ) %>% 
  hideGroup("In-situ") %>%
  addLegend(pal = pal2, values = c(-100, 150), opacity = 1, 
            title = "01 April SWE</br>2024 anomaly",
            position = "bottomright",
            labFormat = labelFormat(transform = \(x) sort(x, decreasing = TRUE),
                                    suffix = " %"))
  
```

# About

## Column 

### Notes

Note that ERA5-Land has a simplistic snow scheme and coarse resolution for complex mountain terrain, but nonetheless shows good accuracy for large-scale SWE monitoring (see references). Positive SWE anomalies capped to a maximum of 150%. 

SWE values (and consequently anomalies) for glaciated areas, such as in the Arctic, might not be reliable.

### Version history

v0.2 (2024-06-14)

- added more references and notes in About page
- added in-situ anomalies to main map, layer turned off by default (only North America, see also [here](https://github.com/jb-smsc/snotel_anoms))

v0.1 (2024-05-29)

- first online version

## Column 

### About


This is a contribution to the IACS joint body on mountain snow cover, working group 2.

Peak snow water equivalent (SWE) extracted from ERA5-Land for the GMBA mountain inventory v2.0. Peak SWE is taken as for Apr 1. Maps show anomalies of the 2023-2024 season relative to the climatology 1991-2020.

GBMA polygons have been simplified by a factor of 5 for simpler and faster online accessibility.

ERA5-Land processing by S. Gascoin ([sgascoin](https://github.com/sgascoin)), in-situ data by C. Mortimer ([mortc](https://github.com/mortc)), dashboard by M. Matiu ([mitmat](https://github.com/mitmat))




### References 

Snethlage, M.A., Geschke, J., Spehn, E.M., Ranipeta, A., Yoccoz, N.G., Körner, Ch., Jetz, W., Fischer, M. & Urbach, D. A hierarchical inventory of the world’s mountains for global comparative mountain science. Nature Scientific Data. https://doi.org/10.1038/s41597-022-01256-y (2022).
Dataset

Snethlage, M.A., Geschke, J., Spehn, E.M., Ranipeta, A., Yoccoz, N.G., Körner, Ch., Jetz, W., Fischer, M. & Urbach, D. GMBA Mountain Inventory v2. GMBA-EarthEnv. https://doi.org/10.48601/earthenv-t9k2-1407 (2022).

Muñoz Sabater, J. (2019): ERA5-Land hourly data from 1950 to present. Copernicus Climate Change Service (C3S) Climate Data Store (CDS). https://doi.org/10.24381/cds.e2161bac. Date Accessed 04-24-2024. 

Sourp, L., Gascoin, S., Jarlan, L., Pedinotti, V., Bormann, K. J., and Baba, M. W.: Evaluation of high resolution snowpack simulations from global datasets and comparison with Sentinel-1 snow depth retrievals in the Sierra Nevada, USA, EGUsphere [preprint], https://doi.org/10.5194/egusphere-2024-791, 2024. 

Mudryk, L., Mortimer, C., Derksen, C., Elias Chereque, A., and Kushner, P.: Benchmarking of SWE products based on outcomes of the SnowPEx+ Intercomparison Project, EGUsphere [preprint], https://doi.org/10.5194/egusphere-2023-3014, 2024. 

